---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# OpenLand

<!-- badges: start -->
[![Travis build status](https://travis-ci.org/tredgi/OpenLand.svg?branch=master)](https://travis-ci.org/tredgi/OpenLand)
<!-- badges: end -->

OpenLand is an open-source R package for the analysis of Land Use
and Land Cover (LULC)  time series. It includes support for loading multitemporal
gridded datasets; extraction of LULC transition matrices; determination of
net and loss changes; performing a complete LULC change intensity analysis; and
visualization of LULC change and outcomes of intensity analysis using various 2D plots.

## Installation

You can install the released version of OpenLand from GitHub repository with:

```{r installation, eval = FALSE}
devtools::install_github("tredgi/OpenLand")
```


## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(OpenLand)
## basic example code
```

