---
title: "Quick Intro for OpenLand Package"
output: rmarkdown::html_vignette
bibliography: papers_OpenLand.bib
vignette: >
  %\VignetteIndexEntry{Quick Intro for OpenLand Package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<style>
body {
text-align: justify}
</style>

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

**This is a Vignette on how to use the OpenLand package into Land Use and Land Cover (LULC)
data manipulation and producing visualisations.**

## Description of the tools

OpenLand is an open-source R package for the analysis of land use and land cover
time series. It includes support for loading spatiotemporal data `raster map`; 
simulating landscapes with some degree of spatial correlation; extraction of
land use transition matrices; computing of net change and gross changes in a given 
interval; performing a complete intensity analysis based on [@Aldwaik2012]. It provides
methods for visualizations of land use and land cover change like 
the outcomes of intensity analysis, as other graphs like sankey diagram 
(onestep or multistep) and chord diagram displaying the inter-relationships 
between the land uses.


## Data source
The datasets we are going to use for this experiment is from the SOSPantanal...

This is the steps in using OpenLand for LULCC data Analysis

We first need to load these package programs firs:

```{r setup}
library(OpenLand)
```







# References
