---
title: "Quick Intro for OpenLand Package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Quick Intro for OpenLand Package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<style>
body {
text-align: justify}
</style>

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(OpenLand)
```

# Description of the tools

OpenLand is an open-source R package for the analysis of land use and land cover
time series. It includes support for loading spatiotemporal data `raster map`; 
simulating landscapes with some degree of spatial correlation; extraction of
land use transition matrices; computing of net change and gross changes in a given 
interval; performing a complete intensity analysis based on Pontius [@Pontius20011] . It provides
methods for visualizations of land use and land cover change like 
the outcomes of intensity analysis, as other graphs like sankey diagram 
(onestep or multistep) and chord diagram displaying the inter-relationships 
between the land uses.


