---
title: "Quick intro for OpenLand Package"
output: rmarkdown::html_vignette
bibliography: papers_OpenLand.bib
vignette: >
  %\VignetteIndexEntry{Quick Intro for OpenLand Package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<style>
body {
text-align: justify}
</style>

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(OpenLand)
```

```{r include=FALSE}

caseStudy <- SL_2002_2014

caseStudy$tb_legend$className <- factor(c("Ap", "FF", "SA", "SG", "aa", "SF", 
                                          "Agua", "Iu", "Ac", "R", "Im"),
                              levels = c("FF", "SF", "SA", "SG", "aa", "Ap", 
                                         "Ac", "Im", "Iu", "Agua", "R")
  )
# add the color by the same order of the legend,
# it can be the colour name (eg. "white") or the HEX value (eg. #FFFFFF)
caseStudy$tb_legend$color <- c("#CDB79E", "#006400", "#66CD00", "#7FFF00",
                               "#EE2C2C", "#458B00", "#0000CD", "#6B6B6B",
                               "#CD950C", "#6B8E23", "#CD00CD")

```


----
**This is a Vignette on how to use the OpenLand package into Land Use and Land Cover (LULC)
data manipulation and producing visualisations.**

## Description of the tools

OpenLand is an open-source R package for the analysis of land use and land cover
time series. It includes support for loading spatiotemporal data `raster map`; 
simulating landscapes with some degree of spatial correlation; extraction of
land use transition matrices; computing of net change and gross changes in a given 
interval; performing a complete intensity analysis based on [@Aldwaik2012]. It provides
methods for visualizations of land use and land cover change like 
the outcomes of intensity analysis, as other graphs like sankey diagram 
(onestep or multistep) and chord diagram displaying the inter-relationships 
between the land uses.


## Example with (Data source)
The data sets we are going to use for this experiment is from the SOSPantanal...

This is the steps in using OpenLand for LULCC data Analysis

We first need to load these package programs firs:



## The contingency table in a long format

This function calculate a contingency table between a series of points in time, Time$_{t}$ and Time$_{t+1}$
and give the number of elements (C~_tij_~) that transits from category _i_ to category _j_ during interval
this interval of time.

A prototype of this table is as following:

|[Y~_t_~, Y~_t+1_~] | Category~_i_~ | Category~_j_~ |C~_tij_~(km^2^)|Y~_t+1_~ - Y~_t_~|C~_tij_~(pixel)|Y~_t_~|Y~_t+1_~|
|:-----------:|:----------:|:----------:|:--------------:|:-----------:|:--------------:|:----:|:----:|
|Period of analysis from time point _t_ to time point _t+1_|A category at interval’s initial time point|A category at  interval’s final time point|Number of elements in km^2^ that transits from category _i_ to category _j_|Interval in years between time point _t_ and time point _t+1_|Number of elements in pixel that transits from category _i_ to category _j_|Initial Year of the interval| Final Year of the interval|
|Period|From|To|km2|interval|QtPixel|yearFrom|yearTo|


For our study with the function `contingenceTable(input_raster = SaoLourencoBasin, pixelresolution = 30)` area gives this contingency table:

(@) The first object (``r names(caseStudy[1])``) returned from this function is a contingency table with all the points of time analysed.

```{r}

caseStudy$lulc_Multistep

```

```{r eval=FALSE, include=FALSE}
knitr::kable(
caseStudy$lulc_Mulstistep[1:10, ]
)
```

(@) The second object (``r names(caseStudy[2])``) returned from this function is a table containing the class name 
associate with every pixel value and a color to be associate (important for the plots).

```{r}

caseStudy$lulc_Onestep

```
> This object come with aleatory values in the ``r names(caseStudy$tb_legend[2])`` 
and the ``r names(caseStudy$tb_legend[3])`` columns. These values must be edited 
to the real values corresponding the data to be analysed. `caseStudy$tb_legend$className <- factor(c("Ap", "FF", "SA", "SG", "aa", "SF", "Agua", "Iu", "Ac", "R", "Im"), levels = c("FF", "SF", "SA", "SG", "aa", "Ap", "Ac", "Im", "Iu", "Agua", "R"))`

```{r}
# This first command edit the `className` column with the values that represents the dataset
# caseStudy$tb_legend$className <- factor(c("Ap", "FF", "SA", "SG", "aa", "SF",
#                                               "Agua", "Iu", "Ac", "R", "Im"),
#                                  levels = c("FF", "SF", "SA", "SG", "aa", "Ap",
#                                              "Ac", "Im", "Iu", "Agua", "R"))

# this command edit the `color` column with the colors apropriate with the dataset
# caseStudy$tb_legend$color <- c("#CDB79E", "#006400", "#66CD00","#7FFF00",
#                                   "#EE2C2C", "#458B00", "#0000CD", "#6B6B6B",
#                                   "#CD950C", "#6B8E23", "#CD00CD")
                                  
```


(@) The third object (``r names(caseStudy[3])``) returned from this function is a table containing the total area of study in km^2^ and in pixel unit.

```{r}
caseStudy$tb_legend

```

(@) The fourth object (``r names(caseStudy[4])``) is a contingency table, with the unique difference is the interval the first year is Y~t~ and the final year is Y~T~.

```{r}

caseStudy$totalArea

```
(@) The fifth object (``r names(caseStudy[5])``) is the total interval between the first year (Y~t~) and the last year(Y~T~) in year unit.

```{r}

caseStudy$totalInterval

```



## Intensity Analysis

```{r}
testSL <- intensityanalysis(dataset = caseStudy,
                            class_n = "Ap", class_m = "SG")

# it's returns a list with 10 objects
names(testSL)
```

## The graphics


### Interval Level

```{r, fig.width=7, fig.height=4}

plot(testSL$interval_lvl)

```

### Category Level

* Gain

```{r, fig.width=7, fig.height=4}

plot(testSL$category_lvlGain)

```

* Loss
```{r, fig.width=7, fig.height=4}

plot(testSL$category_lvlLoss)

```

### Transition Level

* Gain
```{r, fig.width=7, fig.height=4}

plot(testSL$transition_lvlGain_n)

```

* Loss
```{r, fig.width=7, fig.height=4}

plot(testSL$transition_lvlLoss_m)

```

## Net and Gross gain and loss

```{r fig.width=7, fig.height=4}

netgrossplot(dataset = caseStudy$lulc_Multistep, # gross gain and loss
             legendtable = caseStudy$tb_legend,
             title = NULL)

```


## Chord Diagram (2002 - 2014)

```{r fig.width=8, fig.height=7}

circlizeplot(dataset = caseStudy$lulc_Onestep,
             legendtable = caseStudy$tb_legend)

```

## Sankey Multi Step

```{r, fig.width=8, fig.height=4}

sankeyLand(dataset = caseStudy$lulc_Multistep,
           legendtable = caseStudy$tb_legend)


```


```
           2002                 2008                2010                  2012                  2014
```

## Sankey One Step

```{r, fig.width=8, fig.height=4}

sankeyLand(dataset = caseStudy$lulc_Onestep,
           legendtable = caseStudy$tb_legend)

```
```
                          2002                                                  2014
```

## An Evolution Bar Plot

```{r, fig.width=7, fig.height=4}

anualplot(dataset = caseStudy$lulc_Multistep, 
          legendtable = caseStudy$tb_legend,
          area_km2 = T)
```



# References
